---
title: "2.03_Flexible_Linear_Models"
output: html_document
---

### Flexible Linear Models:

# Models for complicated study designs

The treatment-contrast parametization
Each coefficient represents the difference from a base condition

An alternative is:
Group-Means parametization 
Y=B1(X1) + B2(X2) + E
B1 = mean in ER-neg
B2 = mean in ER-pos
Tets: B2 - B1 = 0 

No intercept coefficient
Therefore you have to construct contrasts to test for differential expression

# To include a model without an intercept term:
```{r}
group_design <- model.matrix(~0 + Disease, data = pData(sCLLex))
colSums(group_design)
cm <- makeContrasts(Progression = Diseaseprogres. - Diseasestable,
                    levels = group_design)
# With makeContrasts you can make any contrast you'd like by just referring to the column names of the design matrix
cm
```

#  Testing the group-means parametrization
```{r}
fit <- lmFit(sCLLex, group_design)
 
head(fit$coefficients, 3)

fit2 <- contrasts.fit(fit, contrasts = cm)

head(fit2$coefficients, 3)
```

# The parametrization does not change the results
```{r}
#Calculate the t-statistics
fit2 <- eBayes(fit2)
 
# Count the number of differentially expressed genes
results <- decideTests(fit2)
summary(results)
```


## Studies with more than two groups:
3 different types of leukemias: ALL, AML, CML
Bioconductor package: leukemiasEset
Kohlmann et al. 2008, Haferlach et al. 2010

```{r}
BiocManager::install("leukemiasEset")
library("leukemiasEset")

data(leukemiasEset)

library(annotate)
library(hgu95av2.db)
ID_1 <- featureNames(leukemiasEset)
Symbol_1 <- getSYMBOL(ID_1,"hgu95av2.db")
fData(leukemiasEset) <- data.frame(Symbol=Symbol_1)


# Expression matrix
a <- exprs(leukemiasEset)

# Feature data
b <- fData(leukemiasEset)

# Phenotype data
c <- pData(leukemiasEset)


dim(leukemiasEset)
table(pData(leukemiasEset)[, "LeukemiaType"])
```

# Build a group-means model for 3 groups
Y=B1(X1) + B2(X2) + B3(X3) + E
B1 = mean in ER-neg
B2 = mean in ER-pos
Tests:
AML v. ALL: B2 - B1 = 0
CML v. ALL: B3 - B1 = 0
CML v. AML: B3 - B2 = 0

The process is identical to the 2 group case

```{r}
design <- model.matrix(~0 + LeukemiaType, data = pData(leukemiasEset))

head(design, 3)

colSums(design)
```

# Make contrasts for 3 groups:
```{r}
library(limma)
cm <- makeContrasts(AMLvALL = LeukemiaTypeAML - LeukemiaTypeALL,
                    CMLvALL = LeukemiaTypeCML - LeukemiaTypeALL,
                    CMLvAML = LeukemiaTypeCML - LeukemiaTypeAML,
                    levels = design)

cm
```

# Testing 3 groups
```{r}
# Fit coefficients
fit <- lmFit(leukemiasEset, design)

# Fit contrasts
fit2 <- contrasts.fit(fit, contrasts = cm)

# Calculate t-statistics
fit2 <- eBayes(fit2)

# Summarize results
results <- decideTests(fit2)
summary(results)
```

CML and ALL have the biggest difference



## Factorial experimental design:

How to model studies with a factorial experimental design
2x2 design:
Example: Study effect of low temperature in plants: 2 types of plants
col, vte2 (3 replicates each)
Exposed to normal and low temperature

# Model for 2x2 factorial:
Create a coefficient for each group:
Y=B1(X1) + B2(X2) + B3(X3) + B4(X4) + E
B1 = mean in col plants at low temperature
B2 = mean in col plants at low temperature
B3 = mean in vte2 plants at low temperature
B4 = mean in vte2 at normal temperature

# Need to create a new group that captures both variables:
group <- with(pData(eset), paste(type, temp, sep = "."))
# Convert to a factor
group <- factor(group)

design <- model.matrix(~0 + group)
# Replace column names with factor levels
colnames(design) <- levels(group)

head(design, 3)
  col.low col.normal vte2.low vte2.normal
1       0          1        0           0
2       0          1        0           0
3       0          1        0           0

colSums(design)
 
col.low  col.normal    vte2.low vte2.normal 
3           3           3           3

# Make Contrasts for a 2x2 factorial:
      B1 B2 B3 B4
type	col	col	vte2	vte2
temp	low	normal	low	normal

Contrasts we're interested in:
Differences of type in normal temp
Differences of type in low temp: 
Differences of temp in vte2
Differences of temp in col type:
Differences of temp between col and vte2 type: (B3-B4) - (B1-B2) = 0
# This is traditionally referred to as an interaction effect - how does the response to temp differ between the two types of plants

library(limma)
cm <- makeContrasts(type_normal = vte2.normal - col.normal,
                    type_low = vte2.low - col.low,
                    temp_vte2 = vte2.low - vte2.normal,
                    temp_col = col.low - col.normal,
                    interaction = (vte2.low - vte2.normal) -
                                  (col.low - col.normal),
                    levels = design)
                    
Contrasts
Levels        type_normal type_low temp_vte2 temp_col interaction
  col.low               0       -1         0        1          -1
  col.normal           -1        0         0       -1           1
  vte2.low              0        1         1        0           1
  vte2.normal           1        0        -1        0          -1
  

# Fit coefficients
fit <- lmFit(eset, design)

# Fit contrasts
fit2 <- contrasts.fit(fit, contrasts = cm)

# Calculate t-statistics
fit2 <- eBayes(fit2)

# Summarize results
results <- decideTests(fit2)
summary(results)  

   type_normal type_low temp_vte2 temp_col interaction
-1           0      466      1635     1885         128
0        11871    10915      7635     6989       11640
1            0      490      2601     2997         103
# Conclusion: The plants have similar responses at normal temps but differ at low temps

# Exercises:
2x2 design to study effect of drought in trees:
2 types of Populus: DN34, NM6
2 water conditions: normal, drought
Wilkins et al. 2009
 
dim(eset)
 
Features  Samples 
   16172       12
 
table(pData(eset)[, c("type", "water")])
 
water
type   drought normal
  dn34       3      3
  nm6        3      3
  


