---
title: "3.03_Normalizing_and_Filtering"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Normalizing and Filtering:

# Pre-processing steps:
For technology specific pre-processing steps,
Check the bioconductor workflow course in datacamp!!!

The following should help with most data types:
log transform
quantile normalize
filter

# Plot distribition of each sample
plotDensities(eset, legend=FALSE)

log transform - increase the distance between small measurements, and decrease the distance between the large measurements

exprs(eset) <- log(exprs(eset))
plotDenstities(eset, legend=FALSE)

To reduce the effects of large scale differences due to technical artifacts, one can use Qauntile normalization
e

# Quantile normalize
```{r}
sCLLex_norm<-sCLLex 
exprs(sCLLex_norm) <- normalizeBetweenArrays(exprs(sCLLex))

plotDensities(sCLLex, legend = FALSE)
plotDensities(sCLLex_norm, legend = FALSE)
```
This will ensure that all samples have the same distribution

# View the normalized data
plotDensities(eset, legend = FALSE)
abline(v = 5)
```{r}
plotDensities(sCLLex_norm, legend = FALSE)
abline(v = 5)
# Create logical vector
keep <- rowMeans(exprs(sCLLex_norm)) > 5
# Filter the genes
sCLLex_norm <- sCLLex_norm[keep, ]
plotDensities(sCLLex_norm, legend = FALSE)
```

s


