---
title: "3.05_Accounting_For_Technical_Batch_Effects"
output: html_document
---

### Accounting for technical batch effects

# Dimension reduction techniques:

* Principal Components Analysis (PCA)
* MultiDimensional Scaling (MDS)
* Identify the largest sources of variation in a data set

Are the largest sources of variation correlated with the variables of interest or technical batch effects?

# plotMDS
```{r}
library(limma)
plotMDS(sCLLex_norm, labels = pData(sCLLex_norm)[, "Disease"], gene.selection = "common")
```

# Remove batch effects:
```{r}
exprs(eset) <- removeBatchEffect(eset, batch = pData(eset)[, "batch"],
                                 covariates = pData(eset)[, "rin"])
```


## Example dataset:
Olfactory stem cells
7 treatments, 4 batches
Bioconductor package: HarmanData, Harman
Osmond-McLeod et al. 2013, Oytam et al. 2016

```{r}
BiocManager::install("HarmanData")
library("HarmanData")

BiocManager::install("Harman")
library("Harman")

data(OLF)

OLF_eset <- ExpressionSet(assayData = as.matrix(olf.data),
                      phenoData = AnnotatedDataFrame(olf.info))

plotMDS(OLF_eset, labels = pData(OLF_eset)[, "Treatment"], gene.selection = "common")

plotMDS(OLF_eset, labels = pData(OLF_eset)[, "Batch"], gene.selection = "common")


# remove batch effect:
OLF_eset_batchcorr <- OLF_eset
exprs(OLF_eset_batchcorr) <- removeBatchEffect(OLF_eset_batchcorr, batch = pData(OLF_eset_batchcorr)[, "Batch"])

plotMDS(OLF_eset_batchcorr, labels = pData(OLF_eset_batchcorr)[, "Treatment"], gene.selection = "common")

plotMDS(OLF_eset_batchcorr, labels = pData(OLF_eset_batchcorr)[, "Batch"], gene.selection = "common")
```


