---
title: "3.10_Visualization_and_Enrichment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Visualizing Results:

# Re-load leukemia data:
```{r}

ID_2 <- featureNames(sCLLex)
Symbol_2 <- getEG(ID_2,"hgu95av2.db")
Symbol <- getSYMBOL(ID,"hgu95av2.db")
fData(sCLLex) <- data.frame(ENTREZ=Symbol_2,Symbol=Symbol)

group_design <- model.matrix(~0 + Disease, data = pData(sCLLex))
colSums(group_design)
cm <- makeContrasts(Progression = Diseaseprogres. - Diseasestable,
                    levels = group_design)
# With makeContrasts you can make any contrast you'd like by just referring to the column names of the design matrix
cm
```

#  Testing the group-means parametrization
```{r}
fit <- lmFit(sCLLex, group_design)
 
head(fit$coefficients, 3)

fit2 <- contrasts.fit(fit, contrasts = cm)

head(fit2$coefficients, 3)
```

# The parametrization does not change the results
```{r}
#Calculate the t-statistics
fit2 <- eBayes(fit2)
 
# Count the number of differentially expressed genes
results <- decideTests(fit2)
summary(results)
```



# Return the most differentially expressed genes:
```{r}
topTable(fit2, number=3)
```

B = log odds (another way of telling if a gene is DE'd)

# Obtain all DEG results:
```{r}
stats <- topTable(fit2, number=nrow(fit2), sort.by="none")
# by not sorting, we'll be able to combine the results with the input data
```


# Histogram of p-values:
Under the null hypothesis of no DE, we'd expect the p-values to be uniformaly distributed
```{r}
hist(stats[,"P.Value"])
```

If you don't get a flat plot, or a plot with many small p-values, you need to check your code and see if your contrasts were selected properly.


# Volcano plots:
```{r}
volcanoplot(fit2, highlight = 5, names = fit2$genes[,"Symbol"])

```

# Enrichment Testing:
Databases: KEGG, 

Using Fisher's exact test for enrichment testing.

you need to 

You need to use ENTREZID:
```{r}
head(fit2$genes, 3)

ENTREZ <- fit2$genes[, "ENTREZ"]

enrich_kegg <- kegga(fit2, geneid = ENTREZ, species = "Hs")

topKEGG(enrich_kegg, number = 3)
enrich_go <- goana(fit2, geneid = ENTREZ, species = "Hs")
topGO(enrich_go, ontology = "BP", number = 3)
```








