---
title: "4.03_Pre-Process-Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Putting it all together: pre-process the data

## Doxorubicin-induced cardiotoxicity study:

2x2 design to study mechanism of doxo
2 genotypes: wild-type (WT), Top2b null (top2b)
2 treatments: PBS (pbs), doxorubicin (dox)
Zhang et al. 2012

Is Top2b required for doxorubicin induced toxocity


dim(eset)
 
Features  Samples 
   29532       12
 
table(pData(eset)[, c("genotype", "treatment")])
 
treatment
genotype dox pbs
   top2b   3   3
   wt      3   3
   
```{r}
plotDensities(eset, group = pData(eset)[, "genotype"], legend = "topright")

# Create a new ExpressionSet to store the processed data
eset <- eset_raw
# Log transform
exprs(eset) <- log(exprs(eset))
plotDensities(eset,  group = pData(eset)[,"genotype"], legend = "topright")
# Quantile normalize
exprs(eset) <- normalizeBetweenArrays(exprs(eset))
plotDensities(eset,  group = pData(eset)[,"genotype"], legend = "topright")
# Determine the genes with mean expression level greater than 0
keep <- rowMeans(exprs(eset)) > 0
sum(keep)
# Filter the genes
eset <- eset[keep,]
plotDensities(eset, group = pData(eset)[,"genotype"], legend = "topright")
```
   
Log transform
Quantile normalize
Filtering

```{r}
# Find the row which contains Top2b expression data
top2b <- which(fData(eset)[,"symbol"] == "Top2b")

# Plot Top2b expression versus genotype
boxplot(exprs(eset)[top2b, ] ~ pData(eset)[,"genotype"],
        main = fData(eset)[top2b, ])
```

# Check sources of variation
Principal component analysis
limma function plotMDS


```{r}
# Plot principal components labeled by genotype
plotMDS(eset, labels = pData(eset)[,"genotype"], gene.selection = "common")

# Plot principal components labeled by treatment
plotMDS(eset, labels = pData(eset)[,"treatment"], gene.selection = "common")
```




